---
title: "Homework One"
author: Mahitha Jangeti
date: September 20 2025
output: github_document
---
## Problem 1

## Loading the Dataset
```{r}
library(moderndive)
data("early_january_weather")
```

## Analyzing the data
The early_january_weather data set contains information on the different aspects of weather in the region, reported by EWR, throughout different hours of the day. 
**origin** (weather station)
**temp** (observed temperature in F)
**dewp** (dewpoint in F)
**humid** (relative humidity)
**wind_dir** (wind direction in degrees)
**wind_speed** (speed and gust speed in mph)
**visib** (visibility in miles)
**time_hour** (date and hour of the recording)


The dataset has a total of `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
The mean temperature is `r mean(early_january_weather$temp)`

## Creating a Scatter-plot
```{r}
library(tidyverse)
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
ggsave("HW1_Scatterplot.pdf", height = 4, width = 6)
```

## Problem 2

## Creating a dataframe
```{r}
homework_df = 
  tibble(
  norm_samp = rnorm(10, mean = 0, sd = 1),
  vec_logical = norm_samp > 0,
  vec_char = c("a", "b", "c" , "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "medium", "high",
                        "low", "medium", "high",
                        "low", "medium", "high",
                        "low")),
)
homework_df
```

## Mean of each variable
```{r}
mean(pull(homework_df, norm_samp))
```

```{r}
mean(pull(homework_df, vec_logical))
```

_Mean does not work for vec_char because this value is not numeric or logical_
```{r}
mean(pull(homework_df, vec_char))
```

_Mean does not work for vec_factor because this value is not numeric or logical. Can be converted to find mean_
```{r}
mean(pull(homework_df, vec_factor))
```

## Converting Variables 
```{r results='hide'}
as.numeric(homework_df$vec_logical)
as.numeric(homework_df$vec_factor)
as.numeric(homework_df$vec_char)
```
**Character** variables **cannot** be converted into a numerical value while **factor** and **logical** variables **can be** converted into numerical values as logical variables turned into 1 or 0 indicating TRUE and FALSE respectively. **Factor** can be converted to numeric as they are stored internally as integers. 

